---
title: "stats_ch4"
author: "Sigrid"
date: "18/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
pacman::p_load(tidyverse, rethinking)
```

# 4.7 Practice
*Easy*

4E1. The first line is the likelihood
yi ~ Normal (mu, sigma)

4E2. Two parameters; mu and sigma

4E3. Bayes' theorem

P(yi | mu,sigma) = (y_i ~ Normal(mu, sigma) * mu~Normal(0,10) * sigma~Exponential(1))/sum(all of the above)

4E4. Line number two is the linear model
mu_i = alpha + beta*x_i

4E5. Three parameters; alpha, beta and sigma
y is not a parameter, because it is the data (outcome variable) and mu_i is no longer a parameter, because it is now deterministic (=) and not probabilistic

*Medium*
4M1. For the model definition below, simulate observed heights from the prior (not the posterior).
ğ‘¦ğ‘–âˆ¼Normal(ğœ‡,ğœ)
ğœ‡âˆ¼Normal(0,10)
ğœâˆ¼Uniform(0,10)

```{r}
n <- 1e4

set.seed(4)

sim <- tibble(
  sample_mu = rnorm(n, 0, 10),
  sample_sigma = rexp(n, rate = 1),
  prior_r = rnorm(n, sample_mu, sample_sigma) 
) 

sim %>% ggplot(aes(x = prior_r)) +
  geom_density(fill = 'grey33') +  
  ggtitle("height ~ dnorm(mu, sigma)") +
  theme(panel.grid = element_blank())
```
4M2. Translate the model just above into quap formula
```{r}
q <- quap(
  alist(
    prior_r ~ dnorm(mu, sigma),
    mu ~ dnorm(0,10),
    sigma ~ dunif(0,10),
  ), data = sim
)

```
4M3. Translate the quap model formula below into a mathematical model definition. 

```{r}

```


```{r}
data(Howell1); d <- Howell1


```

wesanderson

RColorBrewer

scale_fill_manual(values = wes_palette("FantasticFox1"))